### 2024-11-29 paper/paper.qmd
paper update

---
title: "Analysis of Shootings and Firearm Discharges in Toronto"
subtitle: "A Comprehensive Examination Reveals Significant Time and Locational Impacts on Incident Severity"
author:
  - name: Xinze Wu
    affiliation: University of Toronto
    email: kerwin.wu@mail.utoronto.ca
thanks: "Code and data are available at: https://github.com/ke3w/Data-Analysis-of-Shootings-and-Firearm-Discharges-in-Toronto."
date: "2024-11-26"
output: 
  pdf_document:
    toc: true
    number_sections: true
abstract: |
  This paper analyzes the impact of time and location on the severity of shootings and firearm discharges in Toronto, utilizing a Gradient Boosting Machine model. The findings highlight significant variations across different times and locations, emphasizing the need for targeted policy interventions. This study not only aids in understanding the dynamics of urban violence but also assists in resource allocation for law enforcement and public safety initiatives.
bibliography: [references.bib]
---

# Introduction

The safety and security of urban areas are of paramount importance to their residents and governance structures. In Toronto, the occurrence of shootings and firearm discharges poses significant challenges to public safety. Understanding the patterns and determinants of these violent events is crucial for developing effective interventions and policies. This paper employs a Gradient Boosting Machine model to analyze how time-related factors and geographic locations influence the severity of shooting incidents in Toronto.

## Estimand

This study aims to estimate the effect of time (day of week, hour, and time range) and location (neighbourhood and police division) on the severity of shootings and firearm discharges in Toronto, measured by a weighted score combining deaths and injuries.

## Model Setup

We utilize a Gradient Boosting Machine model, chosen for its robustness in handling various types of data and its ability to model complex interactions and nonlinear relationships between predictors and the response variable.

## Importance of the Study

The results of this study are crucial for informing public safety strategies and police resource allocation in Toronto, aiming to reduce the incidence and severity of violent firearm-related incidents.

# Data {#sec-data}

Data for this study is sourced from Open Data Toronto: <https://open.toronto.ca/dataset/shootings-firearm-discharges/>, detailing all recorded shootings and firearm discharges within the city limits from 2004 to 2024.

## Data Cleaning

Data cleaning was performed using R[@citeR] packages such as `tidyverse` and `lubridate`. The `occ_date` variable was converted to a numeric format to fit the model requirements, and categorical variables were transformed using appropriate factor conversions.

## Variables

### Outcome Variable

-   **Weighted Score**: Combines the number of deaths and injuries, weighting deaths twice as heavily(weighted_score = death \* 2 + injuries).

### Predictor Variables

-   **Time Factors**:

-   `occ_date`(date): Date of Offence Occurred

-   `occ_dow` (day of the week): Day of the Month Offence Occurred

-   `occ_time_range` (morning, afternoon, evening):Time Range of Day Offence Occurred

-   **Location Factors**:

-   `neighbourhood_158:`Name of Neighborhood using City of Toronto's new 158 neighborhood structure

-   `division`: Police division where offence occurred

## Data visualization

### Weighted Score Over Time

```{r}
#| warning: false
#| message: false
#| label: fig-date
#| fig_caption: true
#| fig-cap: "line graph shows the relationship between weighted score over date"
#| echo: false

ggplot(analysis_data, aes(x = occ_date, y = weighted_score)) +
  geom_line(group = 1, color = "blue") +
  scale_x_date(date_breaks = "1 year", date_labels = "%Y") +
  labs(title = "Weighted Score Over Time", x = "Date", y = "Weighted Score") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1))
```

This plot @fig-date illustrates the variation in the weighted scores of shootings and firearm discharges in Toronto from 2004 to 2025. The weighted score, which combines the number of deaths and injuries with deaths being given a higher weight, is plotted against the date of incidents. Notably, the plot reveals several significant spikes, indicating periods with higher shooting severity. These peaks might correlate with specific events or changes in local circumstances, warranting further investigation. The majority of the time, however, the weighted scores remain relatively low, suggesting sporadic rather than consistent patterns of high-severity incidents. This visualization highlights the dynamic nature of crime severity over time and underscores the importance of continuous monitoring and analysis to understand the underlying trends and triggers.

### Weighted Score by Day of week

```{r}
#| warning: false
#| message: false
#| label: fig-dow
#| fig_caption: true
#| fig-cap: "Bar chart of the weighted score by the day of week"
#| echo: false

ggplot(analysis_data, aes(x = occ_dow, y = weighted_score)) +
  geom_bar(stat = "identity", fill = "steelblue") +
  labs(title = "Weighted Score by Day of Week", x = "Day of Week", y = "Weighted Score") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

The bar chart @fig-dow illustrates the weighted scores of shootings and firearm discharges in Toronto, broken down by day of the week. Notably, the weekend days (Saturday and Sunday) along with Monday, show elevated scores, suggesting a higher incidence of severe incidents during these days. This trend could be attributed to increased social activities and gatherings during the weekend, potentially leading to more conflicts or accidents. Conversely, the mid-week days (Tuesday and Wednesday) exhibit notably lower scores, indicating fewer severe incidents.

### Weighted Score by Day of Year

```{r}
#| warning: false
#| message: false
#| label: fig-doy
#| fig_caption: true
#| fig-cap: "line graph of the weighted score by the day of year"
#| echo: false

ggplot(analysis_data, aes(x = occ_doy, y = weighted_score)) +
  geom_smooth(method = "loess", colour = "darkgreen") +
  labs(title = "Weighted Score by Day of Year", x = "Day of Year", y = "Weighted Score")

```

The line graph @fig-doy visualizes the trend of shootings and firearm discharges in Toronto across different days of the year. The plot shows a clear seasonal pattern, with peaks generally occurring around mid-year and towards the end, particularly noticeable in the increase during summer and winter months. This could correlate with seasonal activities and social behaviors such as holidays and outdoor gatherings that might contribute to increased incidents. The valleys observed during early spring and late fall might reflect quieter periods with fewer such gatherings.

### Weighted Score by Hour of Day

```{r}
#| warning: false
#| message: false
#| label: fig-hour
#| fig_caption: true
#| fig-cap: "Bar chart shows the weighted score by the hour of day"
#| echo: false

ggplot(analysis_data, aes(x = occ_hour, y = weighted_score)) +
  geom_bar(stat = "identity", fill = "red") +
  labs(title = "Weighted Score by Hour of Day", x = "Hour", y = "Weighted Score")

```

The bar graph @fig-hour presents the distribution of shootings and firearm discharges in Toronto by hour. This visualization demonstrates a notable peak in incidents during the very early hours of the day, specifically around midnight, with a dramatic decline shortly after. A secondary, but less intense, peak occurs in the evening hours, spanning from 8 PM to midnight.

This pattern suggests that incidents tend to happen more frequently during late-night hours, possibly linked to social and recreational activities during these times, or reduced visibility and police presence making these hours more conducive to criminal behavior. Conversely, the hours from early morning to mid-afternoon show markedly lower incidents, reflecting perhaps quieter public activity and higher visibility and vigilance during daylight hours.

### Weighted Score by Time Range

```{r}
#| warning: false
#| message: false
#| label: fig-range
#| fig_caption: true
#| fig-cap: "Bar chart shows the weighted score by the time range"
#| echo: false

ggplot(analysis_data, aes(x = occ_time_range, y = weighted_score, fill = occ_time_range)) +
  geom_bar(stat = "identity") +
  labs(title = "Weighted Score by Time Range", x = "Time Range", y = "Weighted Score")

```

This bar graph @fig-range illustrates the severity of shootings and firearm discharges in Toronto across different time range of the day. The time ranges are categorized into morning, afternoon, evening, and night.

The graph reveals a stark increase in incident severity during the night, which towers over the scores recorded for other times of the day. This is followed by the evening, which also shows a considerably high level of incident severity. The overall trend of this graph is highly corresponding to the bar chart of weighted score over hour of the day, basically, this graph provides a more straightforward relationship between the weight score over different periods in a day.

### Weighted Score by Neighborhood

```{r}
#| warning: false
#| message: false
#| label: fig-neighborhood
#| fig_caption: true
#| fig-cap: "Bar chart shows the weighted score by the neighborhood"
#| echo: false

ggplot(analysis_data, aes(x = neighbourhood_158, y = weighted_score, fill = neighbourhood_158)) +
  geom_bar(stat = "identity") +
  labs(title = "Weighted Score by Neighbourhood", x = "Neighbourhood", y = "Weighted Score") +
  theme_minimal() +
  theme(axis.text.x = element_blank())

```

The bar chart @fig-neighborhood displays the severity of shootings and firearm discharges across various neighborhoods in Toronto, segmented by the neighbourhood_158 identifiers. Each bar represents a different neighbourhood, color-coded for visual differentiation and labeled with both the neighbourhood name and a unique identifier number.

This visualization highlights the disparity in incident severity across the city, with some neighbourhoods experiencing significantly higher weighted scores than others. Notably, certain areas show peaks which suggest hotspots of violent incidents. Such patterns are crucial for identifying regions that may require more focused law enforcement and public health interventions to mitigate the impact of firearm-related violence.

### Weighted Score by Division

```{r}
#| warning: false
#| message: false
#| label: fig-division
#| fig_caption: true
#| fig-cap: "Bar chart shows the weighted score by division"
#| echo: false

ggplot(analysis_data, aes(x = division, y = weighted_score, fill = division)) +
  geom_bar(stat = "identity") +
  labs(title = "Weighted Score by Division", x = "Division", y = "Weighted Score") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
```

This bar chart @fig-division illustrates the variation in the severity of shootings and firearm discharges across different police divisions in Toronto. Each division is represented by a unique color and labeled with its respective code (D11 through D55), which simplifies identification and comparison across the chart.

The visualization provides a clear depiction of how incident severity is distributed across the police divisions, with some divisions showing considerably higher weighted scores than others.

# Model Details {#sec-model-details}

## Model Description

The analysis employed a Gradient Boosting Machine (GBM) model using the gbm package in R, designed to predict the weighted scores of shootings and firearm discharges in Toronto based on multiple time-related and geographical predictors. The weighted score is a computed metric that assigns double weight to deaths compared to injuries, highlighting the gravity of incidents resulting in fatalities.

## Model Setup

### Response and Predictor variables

The response variable, weighted_score, combines the effects of deaths and injuries, assigning greater importance to the former. Predictors included:

-   `occ_date`: Numeric representation of the incident date.

-   `occ_dow`: Factor variable indicating the day of the week.

-   `occ_doy`: Numeric day of the year to capture seasonal effects.

-   `occ_time_range`: Categorical variable divided into morning, afternoon, evening, and night.

-   `neighbourhood_158`: Factor variable representing different neighborhoods.

-   `division`: Factor variable representing police divisions.

### Model Configuration

The model was configured with the following parameters to control the complexity and fit of the model:

• Number of Trees: 500, providing sufficient model complexity and accuracy.

• Interaction Depth: 4, allowing interactions among up to four predictors.

• Shrinkage: 0.01, controlling the learning rate to avoid overfitting.

• Cross-validation Folds: 5, used to validate the model internally and optimize parameter selection.

## Model Performance

### Variable Importance
The summary function highlighted the relative importance of each predictor shown in @fig-importance. Notably, the neighborhood variable (neighbourhood_158) dominated the model, indicating significant spatial variation in shooting severities across Toronto. The occ_dow and division showed moderate influence, suggesting that day of the week and police division also contribute to variations in incident severity, albeit to a lesser extent.
```{r}
#| warning: false
#| message: false
#| label: fig-importance
#| fig_caption: true
#| fig-cap: "Bar chart shows the importance of the predictor variables"
#| echo: false

# Set graphical parameters
par(cex.axis = 0.4, las = 1)  # Adjust this value to change the size of the axis labels

# Plot variable importance
summary(model, n.trees = 500, plotit = TRUE)
```
### Model Performance
The model’s accuracy was assessed using the Root Mean Square Error (RMSE), calculated at 2.173830. This metric quantifies the average magnitude of the model’s prediction errors, providing a measure of predictive accuracy.

### Result Interpretation
The results underline the critical influence of geographic location (neighbourhood_158) on the severity of shooting incidents, implying that certain areas are systematically associated with higher severity scores. Temporal factors like the day of the week (occ_dow) and specific police divisions also play roles but are less pronounced compared to spatial factors.

# Results

The model highlighted significant predictors and their relative influence on the severity of shootings. `Division` emerged as a particularly influential predictor, indicating that certain areas within Toronto experience distinctly different levels of shooting severity.

# Discussion

## Implications

Findings from this model underscore the need for targeted interventions in specific neighbourhoods and at times where shootings are most severe.

## Limitations and Future Research

While the model provides valuable insights, it also has limitations due to potential unobserved confounders. Future research should consider additional data sources and alternative modeling approaches.

## Policy Recommendations

Based on the model's findings, policy recommendations are proposed to enhance public safety and optimize police resource allocation.

# Conclusion

This study provides a comprehensive analysis of the factors affecting the severity of shootings in Toronto, offering actionable insights for policymakers and law enforcement agencies.

# References